---
title:  "Revisiting the MiniMUGA annotations"
author: "Karl Broman"
date:   "2020-12-18"
output:
    html_document:
        code_folding: hide
        toc: true
        toc_float: true
        toc_depth: 3
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.width=11, fig.height=6.5,
                      message=FALSE, warning=FALSE)
options(width=110)
```


## Introduction

The annotation file I created for the miniMUGA array (2018-11-08) is at
<https://github.com/kbroman/MUGAarrays/blob/master/UWisc/mini_uwisc_v1.csv>.
We'll call this the UWisc annotation file.

The [miniMUGA paper](https://doi.org/10.1534/genetics.120.303596) has
now been published.
Initially [published at bioRxiv](https://doi.org/10.1101/2020.03.12.989400) on 2020-03-14,
it provides official annotations with the [Supplemental
material](https://doi.org/10.25386/genetics.11971941.v1), as [Table
S2](https://gsajournals.figshare.com/articles/dataset/Supplemental_Material_for_Sigmon_et_al_2020/11971941?file=25117973).
We will call this the UNC annotation file.

My goal here is to compare the two sets of annotations.
The UNC annotation file includes many additional markers which have
been added to the miniMUGA array since my initial analysis. And so
I'll first do a basic comparison of the two files, and then I'll go
about building a revised version of UWisc annotations that incorporates
all of the markers.


### Key findings

```{r load_exec_summary}
file <- "mini_revisited_summary.rds"
if(file.exists(file)) {
    exec_summ <- readRDS(file)
} else {
    message <- "**RE RUN FILE**"
    exec_summ <- list(new_markers=message,
                      no_hit=message,
                      transgene=message,
                      n_diff_chr=message,
                      n_matching_chr=message,
                      n_pos_mismatch=message,
                      n_pos_mismatch_old=message,
                      n_pos_mismatch_new=message)
}
```

- The new UNC annotation file for the miniMUGA array has an
  additional `r exec_summ$new_markers` markers that weren't present in the original
  annotation file I'd studied.

- There are `r exec_summ$no_hit` markers with no perfect blast hit in the mouse genome (assembly
  mm10, GRCm38). This is on top of the `r exec_summ$transgene` transgene markers.

- There are `r exec_summ$n_diff_chr` markers where the UNC annotation file has them on a different
  chromosome than is seen in my blast results. All of these were among
  the `r exec_summ$new_markers` markers added to the array.

- The remaining `r broman::add_commas(exec_summ$n_matching_chr)`
  markers have matching chromosome IDs between the UNC annotation
  file and the blast results,
  but `r exec_summ$n_pos_mismatch` of these have discrepancies in their position.
  Of these, `r exec_summ$n_pos_mismatch_old` were on the previous version of the
  array and are on the mitochondria, while the other
  `r exec_summ$n_pos_mismatch_new` are new.



## Preliminaries

Let's load the two annotation files, plus a few packages.

```{r load_packages_and_annotations}
library(data.table)
library(broman)
library(devtools)

uw <- data.table::fread("../UWisc/mini_uwisc_v1.csv", data.table=FALSE)

unc_url <- "https://gsajournals.figshare.com/ndownloader/files/25117973"
unc_file <- "../UNC/miniMUGA_tableS2.csv"
if(!file.exists(unc_file)) download.file(unc_url, unc_file)
unc <- data.table::fread(unc_file, data.table=FALSE)
```

The UWisc annotations had `r add_commas(nrow(uw))` markers, while the
UNC annotation file has `r add_commas(nrow(unc))`
markers.

`r ifelse(all(uw$marker %in% unc$"Marker name"), "All", "**yes Not all**")`
the markers in the UWisc annotations are in the UNC annotation file,
but then there are `r sum(unc$"Marker name" %nin% uw$marker)`
additional markers in the UNC file, and these are distributed across
most chromosomes.

Here's the chromosome distribution of markers in the UWisc
annotations. I'll replace the missing chromosomes with 0 and replace
`M` with `MT` for the mitochondria, as in the new UNC annotations.

```{r chr_distr_uw}
uw$chr[is.na(uw$chr)] <- 0
uw$chr[uw$chr=="M"] <- "MT"
uw$chr_unc[uw$chr_unc=="M"] <- "MT"
table(factor(uw$chr, levels=c(0:19,"X","Y","PAR", "MT")))
```

And here's the chromosome distribution of markers that are only in the
new annotations.

```{r chr_distr_new}
table(factor(unc$chr[!(unc$"Marker name" %in% uw$marker)],
             levels=c(0:19,"X","Y","PAR","MT")))
```

The additional markers are sprinkled throughout the autosomes and X
chromosome, plus a number on "chr 0", but no new markers on Y, PAR, or
mitochondria.

### Comparison of common markers

Let's focus on the markers in the UWisc annotations, and check that
the genomic positions are the same. I'll first grab the subset of
markers in common and get them in the same order

```{r subset_unc}
rownames(unc) <- unc$"Marker name"
uncsub <- unc[uw$marker,]
```

There are `r sum(uncsub$chr != uw$chr)` markers where the chromosomes
are not the same between the two annotation files. These are exactly
the ones where the UNC file says the marker maps uniquely (the column
`bowtie_unique` is 1) but the UWisc found as not unique (the column
`unique` is FALSE).

```{r check_chr_differences}
stopifnot(
    all( (uncsub$bowtie_unique == 1 & uw$unique==FALSE) ==
         (uncsub$chr != uw$chr) )
         )
```


Let's reduce to the `r add_commas(sum(uw$unique))` markers that the
UWisc annotations says map uniquely in the mouse genome.

```{r reduce_to_uwisc_unique}
uncsub2 <- uncsub[uw$unique,]
uwsub <- uw[uw$unique,]
```

For these markers, the chromosome IDs are the same between the two
files, but the basepair positions differ in 7 cases.

```{r positions_differ}
cbind(uncsub2[,c("chromosome", "position", "strand")],
      uwsub[c("marker", "chr", "bp_mm10", "strand")])[uncsub2$position != uwsub$bp_mm10,]
```

They are all mitochondrial markers, and they differ by one base except
for one case that is off by 50 and says the opposite strand
(marker `r uwsub$marker[abs(uncsub2$position - uwsub$bp_mm10) > 1]`).
This is the only marker where the strands are different.


### Probe sequences

For the markers in common, are the probe sequences the same?
For `r sum(nchar(uw$probe)==49)` markers, the probe sequence contained
the SNP and I had trimmed that off.

For the `r add_commas(sum(uncsub$seqB==""))` markers where there's no
seqB and the probe sequence doesn't contain the SNP, the probes are
identical.

```{r probe_identical_when_missing_seqB}
stopifnot( all(uncsub$seqA[uncsub$seqB==""] == uw$probe[uncsub$seqB==""]) )
```

For the other `r sum(uncsub$seqB != "")` markers, the probe sequence
in the UWisc file deletes the last base, and the first 49 bases are
the same in all cases.

```{r probes_with_seqB}
seqA <- uncsub$seqA[uncsub$seqB != ""]
probe <- uw$probe[uncsub$seqB != ""]
stopifnot( all(substr(seqA, 1, 49) == probe) )
```

### Alleles

Finally, let's verify that the info about SNP alleles are the same.
(Indeed, they are.)

```{r check_snp_alleles}
uncsub$snp <- paste0(uncsub$"reference allele", uncsub$"alternate allele")
stopifnot( all(uncsub$snp ==  uw$snp) )
```

## Revised UWisc annotations

I'm going to go back and re-load the UNC annotations and do all of the
various things to create revised UWisc annotations. I'll follow my
[previous report](https://kbroman.org/MUGAarrays/mini_annotations.html) closely.

```{r load_data}
# UNC file
mini_unc <- unc # from above
rownames(mini_unc) <- mini_unc$"Marker name"

# change a few things:
# Marker name -> marker
# chromosome -> chr
# position -> pos
# MT -> M
colnames(mini_unc)[1:3] <- c("marker", "chr", "pos")
mini_unc$chr[mini_unc$chr=="MT"] <- "M"

# Blast results; keep just the perfect matches(?)
mini_blast <- readRDS("../Blast/results_mini2/mini2_blastn_results.rds")
mini_blast <- mini_blast[mini_blast$tot_mismatch==0,]
```


### Summarize blast hits

I'll start with some code to summarize the blast hits. For each
SNP, I want to count the number of perfect blast hits.
As it turns out, the probes have either 1 or 0 perfect hits, and so
we don't have to do what we did with the MegaMUGA and GigaMUGA arrays,
regarding studying the number of distinct chromosomes hit.

For markers with a unique hit, I'll record the
chromosome, position, and strand.

```{r mini_summarize_blast}
# no. blast hits
mini_tab <- table(mini_blast$query)
mini_nchr <- mini_nhits <- setNames(rep(0, nrow(mini_unc)), rownames(mini_unc))
mini_nhits[names(mini_tab)] <- mini_tab

# chr,pos,strand for the unique ones
mini_blast_uniq <- mini_blast[mini_blast$query %in% names(mini_nhits)[mini_nhits==1],]
mini_blast_chr <- mini_blast_pos <- mini_blast_strand <- setNames(rep(NA, nrow(mini_unc)), rownames(mini_unc))
mini_blast_chr[mini_blast_uniq$query] <- mini_blast_uniq$chr
mini_blast_pos[mini_blast_uniq$query] <- mini_blast_uniq$snp_pos
mini_blast_strand[mini_blast_uniq$query] <- mini_blast_uniq$strand

# probe sequences
mini_blast_probe <- setNames(mini_unc$seqA, mini_unc$Marker)
mini_trim <- (mini_unc$seqB != "")
mini_blast_probe[mini_trim] <- substr(mini_blast_probe[mini_trim], 1, nchar(mini_blast_probe[mini_trim])-1)

# SNP alleles
mini_blast_snp <- paste0(mini_unc$"reference allele", mini_unc$"alternate allele")

# put all of this stuff into a data frame
mini_uwisc <- data.frame(marker=names(mini_nhits),
                         n_blast_hits=mini_nhits,
                         unique=(mini_nhits==1),
                         unmapped=(mini_nhits==0),
                         chr=mini_blast_chr,
                         pos=mini_blast_pos,
                         strand=mini_blast_strand,
                         snp=mini_blast_snp,
                         probe=mini_blast_probe,
                         stringsAsFactors=FALSE)
rownames(mini_uwisc) <- mini_uwisc$marker
```


### Basic blast results

The miniMUGA array contains
`r add_commas(nrow(mini_unc))` markers, of which (according to the
UNC annotations),
`r add_commas(sum(mini_unc$chr %in% (1:19)))` are on autosomes,
`r add_commas(sum(mini_unc$chr=="X"))` are on the X chromosome,
`r add_commas(sum(mini_unc$chr=="Y"))` are on the Y chromosome,
`r add_commas(sum(mini_unc$chr=="M"))` are on the mitochondrial genome,
`r add_commas(sum(mini_unc$chr=="PAR"))` are in the pseudoautosomal region,
and
`r add_commas(sum(mini_unc$chr=="0"))` have chromosome `0` which are
transgene-related markers.

In the blast results, I find that `r add_commas(sum(mini_nhits==1))`
markers have a single, unique hit in the mouse genome (build mm10)
and `r add_commas(sum(mini_nhits==0))` have no perfect hit.
There are
`r ifelse(sum(mini_nhits>1)>0, sum(mini_nhits>1), "no")`
markers with multiple hits.

The markers with no perfect hit include
`r ifelse(all(mini_uwisc$unmapped[mini_unc$chr==0]), "all", sum(mini_uwisc$unmapped[mini_unc$chr==0]))`
of the transgene-related markers, plus another
`r sum(mini_uwisc$unmapped & mini_unc$chr != 0)`
markers that are spread across all autosomes and the X and Y chromosomes.



### Chromosome assignments

If we look at the inferred chromosome assignments of the markers with
a unique hit to the genome, there are
`r sum(mini_unc$chr != mini_uwisc$chr, na.rm=TRUE)`
differences.

This is **many** more discrepancies than we saw in the first version of
the UNC annotation file. Previously, we just saw the
`r numbers[sum(mini_unc$chr=="PAR")]` markers annotated
to be in the pseudoautosomal region (PAR); the blast results match
these to the X chromosome. We still see these, and we'll force them
back to `PAR`.

```{r x_to_par}
mini_uwisc$chr[mini_unc$chr=="PAR"] <- "PAR"
```

But in addition there are another `r sum(mini_unc$chr != mini_uwisc$chr, na.rm=TRUE)`
markers that map to different chromosomes than seen in the UNC annotation
file.

```{r problem_markers}
diff_chr <- mini_uwisc$marker[!is.na(mini_uwisc$chr) & mini_unc$chr != mini_uwisc$chr]
```

In the UNC annotations, these span most autosomes (all except chr 2) plus the
X chromosome, while in the blast results, they are on chr
`r vec2string(sort(as.numeric(unique(mini_uwisc[diff_chr, "chr"]))))`.
And `r ifelse(!any(diff_chr %in% uw$marker), "**none**", sum(diff_chr %in% uw$marker))`
of these markers were in the original version of the miniMUGA array.
They are all among the `r nrow(mini_unc) - nrow(uw)` new markers.

For example, consider marker `r (marker <- "DX1033615090")`, with probe sequence
``r mini_unc[marker, "seqA"]``.
In the UNC annotations, it is on chr `r mini_unc[marker, "chr"]`
at `r add_commas(mini_unc[marker, "pos"])` bp, but in the blast
results, it shows up on chr `r mini_uwisc[marker, "chr"]` at
`r add_commas(mini_uwisc[marker,"pos"])` bp.

As another example, consider marker `r (marker <- "DX1150407521")`, with probe sequence
``r mini_unc[marker, "seqA"]``.
In the UNC annotations, it is on chr `r mini_unc[marker, "chr"]`
at `r add_commas(mini_unc[marker, "pos"])` bp, but in the blast
results, it shows up on chr `r mini_uwisc[marker, "chr"]` at
`r add_commas(mini_uwisc[marker,"pos"])` bp.

If we paste these probe sequences into the [NCBI blast
website](https://blast.ncbi.nlm.nih.gov/Blast.cgi?PAGE_TYPE=BlastSearch&BLAST_SPEC=OGP__10090__9559&LINK_LOC=blasthome),
using the GRCm38 assembly, it confirms my blast results.
Using the GRCm39 assembly, we get the same chromosome assignments but
slightly different positions.


### Positions

For the remaining `r add_commas(sum(!is.na(mini_uwisc$chr) & mini_uwisc$chr != "PAR" & mini_unc$chr == mini_uwisc$chr))`
markers, UNC chromosome assignment and the
blast hit do match, but
`r sum(!is.na(mini_uwisc$chr) & mini_uwisc$chr != "PAR" & mini_unc$chr == mini_uwisc$chr & mini_unc$pos != mini_uwisc$pos)`
of them have some discrepancy in their position.

```{r other_markers}
chr_match <- mini_unc$marker[!is.na(mini_uwisc$chr) & mini_uwisc$chr != "PAR" &
                             mini_unc$chr == mini_uwisc$chr]
stopifnot( all( mini_unc[chr_match, "chr"] == mini_uwisc[chr_match, "chr"] ))
pd <- mini_unc[chr_match, "pos"] - mini_uwisc[chr_match, "pos"]
diff_pos <- chr_match[pd != 0]
result <- cbind(mini_unc[chr_match, c("chr", "pos")], mini_uwisc[chr_match, c("chr", "pos")], pd)[(pd != 0),]
colnames(result) <- c("unc_chr", "unc_pos", "blast_chr", "blast_pos", "pos_diff")
result
```

The `r sum(result$unc_chr=="M")` mitochondrial markers were
all on the original version of the array and are off by mostly 1 bp
(one marker is off by 50). The other
`r sum(result$unc_chr != "M")` markers are new, and are off by quite
varying amounts.

The strand assignments are the same for all `r add_commas(length(chr_match))`
of the markers with matching chromosome assignments, except for the
one mitochondrial marker that is off by 50 bp,
`r rownames(result)[result$blast_chr=="M" & result$pos_diff== -50]`.

```{r verify_strand_assignment}
strand_unc <- setNames(mini_unc[chr_match,"strand"], chr_match)
strand_uwisc <- setNames(mini_uwisc[chr_match,"strand"], chr_match)
strand_uwisc[strand_uwisc=="plus"] <- "+"
strand_uwisc[strand_uwisc=="minus"] <- "-"
stopifnot( sum(strand_unc != strand_uwisc) == 1)
mt_marker <- rownames(result)[result$blast_chr=="M" & result$pos_diff== -50]
stopifnot( strand_unc[mt_marker] != strand_uwisc[mt_marker] )
```



### Markers with the same probes

Not all of the probes are distinct.

```{r dup_probes}
tab_probe <- table(mini_uwisc$probe)
pairs <- names(tab_probe)[tab_probe=="2"]
trios <- names(tab_probe)[tab_probe=="3"]

# names of the markers in the dup pairs
pair_names <- t(sapply(pairs, function(pr) rownames(mini_uwisc)[mini_uwisc$probe == pr]))
rownames(pair_names) <- 1:nrow(pair_names)
pair_chr <- cbind(mini_unc[pair_names[,1],"chr"], mini_unc[pair_names[,2],"chr"])

# pairs from these that are not on chr 0
pairs_not0 <- pair_names[rowSums(pair_chr=="0")<2,]

# verify name mutations
stopifnot( all(pairs_not0[,2] == paste0(pairs_not0[,1], "b") |
    pairs_not0[,2] == paste0("g", pairs_not0[,1]) |
    pairs_not0[,2] == paste0("m", pairs_not0[,1]) |
    pairs_not0[,1] == paste0(pairs_not0[,2], "b") |
    pairs_not0[,1] == paste0("g", pairs_not0[,2]) |
    pairs_not0[,1] == paste0("m", pairs_not0[,2]) |
    pairs_not0[,1] == sub("^D", "S", pairs_not0[,2]) ))
```

There are `r length(pairs)` pairs and `r length(trios)` trios of
markers with identical probe sequences.

The trios of markers with identical probe sequences are all on chr `0`
in the UNC annotation file, for transgene-related markers.

Of the `r length(pairs)` pairs, `r sum(rowSums(pair_chr=="0")==2)` are
both on chromosome `0`.

The other `r length(pairs_not0)` pairs of markers with identical probe
sequences appear to be intended duplicates: they all are such the one
name is the same as the other but with `g` or `m` added to the
beginning, or with `b` added to the end, or in one case `ST` is
replaced by `DT` but the rest of the marker name is the same
(markers `ST2023174930` and `DT2023174930`).

That last example, though, is the one pair where the UNC annotation
file has the markers placed on different chromosomes.

```{r verify_chr}
pairs_not0 <- cbind(pairs_not0, chr1=mini_unc[pairs_not0[,1], "chr"],
                    chr2=mini_unc[pairs_not0[,2], "chr"],
                    pos1=mini_unc[pairs_not0[,1], "pos"],
                    pos2=mini_unc[pairs_not0[,2], "pos"])
stopifnot( sum(pairs_not0[,"chr1"] != pairs_not0[,"chr2"] | pairs_not0[,"pos1"] != pairs_not0[,"pos2"]) == 1)

wh <- pairs_not0[pairs_not0[,"chr1"] != pairs_not0[,"chr2"] | pairs_not0[,"pos1"] != pairs_not0[,"pos2"],1:2]
mini_unc[wh, c("chr", "pos", "strand", "seqA")]
```

The position for `ST2023174930` (chr `r mini_uwisc[wh[1],"chr"]` at
`r add_commas(mini_uwisc[wh[1],"pos"])` bp) looks to be the correct position.


### New annotation file

I now want to save my findings as a new annotation file for the
miniMUGA array. I'll also create a metadata file ("data
dictionary") that explains the columns.

The final annotation files and all source materials are at
<https://github.com/kbroman/MUGAarrays>.


#### Version pre_2

We'll call this version "pre 2". It'll be like the
MegaMUGA and GigaMUGA annotation files I made, but without
`multi`, `n_blast_hits`, or `n_blast_chr`. And I'll include columns `chr_unc`
and `bp_unc`, the chromosome and position in the UNC annotation file.
(This is basically a preliminary file before I go and get the genetic
map positions from the [mouse map converter](http://cgd.jax.org/mousemapconverter/) site.)

```{r write_pre_v2_files}
# order of columns
cols <- c("marker", "chr", "pos", "strand", "snp", "unique",
          "unmapped", "probe", "chr_unc", "pos_unc")
# revised names
cols_new <- c("marker", "chr", "bp_mm10", "strand", "snp", "unique",
              "unmapped", "probe", "chr_unc", "bp_unc")

# MegaMUGA file
mini_file <- "../UWisc/mini_uwisc_pre_v2.csv"

# reorder and rename columns
mini_uwisc <- cbind(mini_uwisc, chr_unc=mini_unc$chr, pos_unc=mini_unc$pos)
mini_uwisc <- mini_uwisc[,cols]
colnames(mini_uwisc) <- cols_new

# reorder rows
mini_uwisc <- mini_uwisc[order(factor(mini_uwisc$chr, levels=c(1:19,"X","Y","PAR","M")),
                               mini_uwisc$bp_mm10,
                               factor(mini_uwisc$chr_unc, levels=c(0:19, "X", "Y", "PAR","M"))),]

# write to CSV file
write.table(mini_uwisc, mini_file, sep=",", quote=FALSE,
            row.names=FALSE, col.names=TRUE)
```

I'll also create a dictionary for the file, which explains what the
columns are.

```{r create_dictionaries}
descriptions <- c("Name of SNP marker",
                  "Chromosome",
                  "Physical position in basepairs for mm10 mouse genome build",
                  "Strand (plus/minus) from which the probe sequence was taken",
                  "SNP alleles as a two-character string",
                  "TRUE indicates that the probe sequence appears exactly once in mm10 mouse genome build",
                  "TRUE indicates that the probe sequence has no perfect match in mm10 mouse genome build",
                  "Probe sequence (49 or 50 bases); the SNP occurs immediately after",
                  "Chromosome in UNC annotation file",
                  "Physical positions in basepairs in the UNC annotation file")

mini_dict_file <- "../UWisc/mini_uwisc_dict_pre_v2.csv"
output <- data.frame(column=cols_new,
                     description=descriptions,
                     stringsAsFactors=FALSE)
write.table(output, mini_dict_file, sep=",", quote=FALSE,
            row.names=FALSE, col.names=TRUE)
```

Finally, I want to make a file that indicates the common markers,
using the SNP positions to determine which ones are identical.
There are a bunch of them, and they all have related names, like the
pairs with identical probe sequences.

```{r write_common_markers_to_file}
# unique marker positions between the arrays
mini_pos <- setNames(paste(mini_uwisc$chr, mini_uwisc$bp_mm10, sep=":"), mini_uwisc$marker)[mini_uwisc$unique]

# find the positions that are in duplicate within the array
tab <- table(mini_pos)
dup <- names(tab)[tab > 1]

# find the corresponding markers
dup_names <- t(sapply(dup, function(d) names(mini_pos)[mini_pos==d]))

pos_spl <- strsplit(rownames(dup_names), ":")

# create data frame with the results
common <- data.frame(chr=sapply(pos_spl, "[", 1),
                     bp_mm10=sapply(pos_spl, "[", 2),
                     marker1=dup_names[,1],
                     marker2=dup_names[,2],
                     stringsAsFactors=FALSE)
# reorder by genomic position
common <- common[order(factor(common$chr, c(1:19,"X","Y","M")), common$bp_mm10),]

# write to a CSV file
write.table(common, "../UWisc/mini_commonmark_uwisc_v2.csv",
            sep=",", quote=FALSE, row.names=FALSE, col.names=TRUE)

# data dictionary
common_cols <- colnames(common)
common_descriptions <- c("Chromosome ID",
                         "Physical position in basepairs for mm10 mouse genome build",
                         "A miniMUGA marker at that position",
                         "A second miniMUGA marker at that position")
common_dict <- data.frame(column=common_cols,
                          description=common_descriptions,
                          stringsAsFactors=FALSE)
# write to file
write.table(common_dict, "../UWisc/mini_commonmark_uwisc_dict_v2.csv",
            sep=",", quote=FALSE, row.names=FALSE, col.names=TRUE)
```



```{r write_bp_to_file}
# write just chr, bp to files, for use with mouse map converter
# (want to get interpolated cM positions from the Cox and G2F1 maps)
write.table(mini_uwisc[!is.na(mini_uwisc$chr) & mini_uwisc$chr %in% c(1:19,"X"), c("chr", "bp_mm10")],
            "../GenMaps/mini2_bp.txt", sep=" ", quote=FALSE,
            row.names=FALSE, col.names=FALSE)
```



#### Version 2, with genetic maps

I used the [mouse map
converter](http://cgd.jax.org/mousemapconverter/) to convert the mm10
basepair positions of the autosome and X chromosome markers to
sex-averaged cM from the [Cox et al.](https://doi.org/10.1534/genetics.109.105486)
and [Liu et al. (aka G2F1)](https://doi.org/10.1534/genetics.114.161653) genetic maps.

```{r load_genetic_maps}
mini_cox <- data.table::fread("../GenMaps/mini2_cox.txt", header=FALSE, data.table=FALSE)
mini_g2f1 <- data.table::fread("../GenMaps/mini2_g2f1.txt", header=FALSE, data.table=FALSE)

# verify stuff
stopifnot( all(mini_cox[,1] == mini_uwisc$chr[1:nrow(mini_cox)]) )
stopifnot( all(mini_cox[,2] == mini_uwisc$bp_mm10[1:nrow(mini_cox)]) )
stopifnot( all(mini_g2f1[,1] == mini_uwisc$chr[1:nrow(mini_g2f1)]) )
stopifnot( all(mini_g2f1[,2] == mini_uwisc$bp_mm10[1:nrow(mini_g2f1)]) )
```

The G2F1 genetic maps include positions < 0 cM. Seems perfectly okay,
but I'm going to shift the maps, using the overall cM:Mbp ratio on
each chromosome, to make 3 Mbp (the conventional start position for
the mouse genome builds) equal to 0 cM.

```{r shift_g2f1_maps}
shifts <- read.csv("../GenMaps/g2f1_shift.csv",
                   stringsAsFactors=FALSE)
shifts <- setNames(shifts[,2], shifts[,1])

for(chr in c(1:19,"X")) {
    mini_g2f1[mini_g2f1[,1]==chr,5] <- mini_g2f1[mini_g2f1[,1]==chr,5] + shifts[chr]
}
```

```{r incorporate_genetic_maps}
nas <- rep(NA, nrow(mini_uwisc)-nrow(mini_cox))
mini_uwisc <- cbind(mini_uwisc,
                  cM_cox=c(mini_cox[,5], nas),
                  cM_g2f1=c(mini_g2f1[,5], nas))
```

Now I can write the new annotation files.

```{r write_v2_files}
cols_new <- c(cols_new[1:3], "cM_cox", "cM_g2f1", cols_new[-(1:3)])
mini_uwisc <- mini_uwisc[,cols_new]

# write miniMUGA file
mini_file <- "../UWisc/mini_uwisc_v2.csv"
write.table(mini_uwisc, mini_file, sep=",", quote=FALSE,
            row.names=FALSE, col.names=TRUE)
```

And finally, the new data dictionary files.

```{r dict_files_v2}
descriptions <- c(descriptions[1:3],
                  "Sex-averaged cM positions from Cox et al. https://doi.org/10.1534/genetics.109.105486",
                  paste("Sex-averaged cM positions from Liu et al.",
                        "https://doi.org/10.1534/genetics.114.161653",
                        "(shifted to avoid cM positions < 0)"),
                  descriptions[-(1:3)])

mini_dict_file <- "../UWisc/mini_uwisc_dict_v2.csv"
output <- data.frame(column=cols_new,
                     description=descriptions,
                     stringsAsFactors=FALSE)
write.table(output, mini_dict_file, sep=",", quote=FALSE,
            row.names=FALSE, col.names=TRUE)
```



## Session info

```{r construct_exec_summary}
file <- "mini_revisited_summary.rds"
exec_summ <- list(new_markers=nrow(mini_unc)-nrow(uw),
                  no_hit=sum(!mini_uwisc$unique) - sum(mini_unc$chr==0),
                  transgene=sum(mini_unc$chr==0),
                  n_diff_chr=length(diff_chr),
                  n_matching_chr=sum(mini_uwisc$chr != "PAR" & mini_uwisc$chr == mini_uwisc$chr_unc, na.rm=TRUE),
                  n_pos_mismatch=length(diff_pos),
                  n_pos_mismatch_old = sum(diff_pos %in% uw$marker),
                  n_pos_mismatch_new = sum(!(diff_pos %in% uw$marker)))
saveRDS(exec_summ, file)
```

Here are the versions of R and R packages that I am using.

```{r session_info}
devtools::session_info()
```
